COG-ED Manuscript

### 2022-01-10 06_Paper/COG-ED/Manuscript.Rmd
changed some tenses and updated order of methods

---
title             : "When easy is not preferred: An effort discounting paradigm for estimating subjective values of tasks"
shorttitle        : "The CERED paradigm for estimating subjective values"
author: 
  - name          : "Josephine Zerna"
    affiliation   : "1,*"
    corresponding : yes
    address       : "Zellescher Weg 17, 01069 Dresden, Germany"
    email         : "josephine.zerna@tu-dresden.de"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Methodology
      - Formal analysis
      - Investigation
      - Project administration
      - Software
      - Visualization
      - Writing - original draft preparation
      - Writing - review & editing
  - name          : "Christoph Scheffel"
    affiliation   : "1,*"
    role:
      - Conceptualization
      - Methodology
      - Funding acquisition
      - Investigation
      - Project administration
      - Software
      - Writing - review & editing
  - name          : "Corinna Kührt"
    affiliation   : "1"
    role:
      - Formal analysis
      - Writing - original draft preparation
      - Writing - review & editing
      - Visualization
affiliation:
  - id            : "1"
    institution   : "Faculty of Psychology, Technische Universität Dresden, 01069 Dresden, Germany"
authornote: |
  * Shared first authorship
abstract: |
  When individuals set goals, they consider the subjective value (SV) of both the anticipated reward and the required effort, a trade-off that is of great interest to psychological research.
  However, the SV of effort is highly individual, and previous quantification approaches have had two crucial limitations: They presumed a unidirectional relationship between objective task load and the SV of effort, and as a consequence, they could only quantify the SVs of tasks with an objective order of task load.
  One of these approaches is the Cognitive Effort Discounting paradigm by Westbrook et al. (2013).
  We aim to replicate their analysis with our adaptation, the Cognitive and Emotion Regulation Effort Discounting (CERED) paradigm.
  We argue that the CERED paradigm allows two crucial things: Quantifying SVs without assuming that the easiest level is preferred, and quantifying SVs for tasks with no objective order of task load.

@@

The paradigm was written and presented using *Psychopy* [@Peirce2019]. We used *R Studio* [@RCT2020; @RStudioTeam2020] with the main packages *afex* [@Singmann2021] and *bayestestR* [@Makowski2019] for all our analyses.
The R Markdown file used to analyze the data and write this document is freely available at github.com/ChScheffel/CERED.
The data can be downloaded from osf.io/vnj8x/.

## Questionnaires

To assess perceived task load, we will use the 6-item NASA Task Load Index (NASA-TLX) [@Hart1988], where participants evaluate their subjective perception of mental load, physical load, effort, frustration, performance, and time pressure during the task on a 20-point scale.
NFC will be assessed using the 16-item short form of the Need for Cognition Scale [@Bless1994; @Cacioppo1984].
Responses to each item (e.g., "Thinking is not my idea of fun", recoded) will be recorded on a 7-point Likert scale.
The NFC scale shows comparably high internal consistency [Cronbach's $\alpha$ \> .80; @Bless1994; @Fleischhauer2010].
Several other personality questionnaires will be used in this study but are the topic of the Registered Report for the second lab session.
A full list of measures can be found at osf.io/vnj8x/.

## Procedure

Healthy participants aged 18 to 30 years will be recruited using the software *ORSEE* [@Greiner2015].
Participants will fill out the personality questionnaires online and then visit the lab for two sessions one week apart.
In the first session, participants provide informed consent and demographic data before completing the computer-based paradigm.
The paradigm starts with the n-back levels one to four, presented sequentially with two runs per level, consisting of 64 consonants (16 targets, 48 non-targets) per run.
The levels are referred to by color (1-back black, 2-back red, 3-back blue, 4-back green) to avoid anchor effects in the ED procedure.
After each level, participants fill out the NASA-TLX on a tablet.
Then, they complete the ED procedure on screen, where each possible pairing of the four n-back levels is presented in a randomized order.
Participants are instructed to decide as realistically as possible, because one of their choices from the last iteration steps will be randomly chosen for one final run of n-back.
This is only done to incentivise truthful behavior in the ED procedure, so the n-back data of this part will not be analyzed.
The second session consists of an emotion regulation task with negative pictures and the instruction to suppress facial reactions, detach cognitively from the picture content, and distract oneself, respectively.
The paradigm follows the same structure of task and ED procedure, but participants can decide which strategy they want to reapply in the last block.
Participants will receive 30€ in total or course credit for participation.
Study data will be collected and managed using REDCap electronic data capture tools hosted at Technische Universität Dresden [@Harris2009; @Harris2019].
The sample of the pilot study consisted of $N=$ `r length(datalist_ED)` participants (`r round((table(data_quest$gender)[2]/length(datalist_ED))*100, digits = 1)`% female, $M=$`r round(mean(data_quest$age, na.rm = TRUE), digits = 1)`($SD=$`r round(sd(data_quest$age, na.rm = TRUE), digits = 1)`) years old).
In the pilot study, the online questionnaires were completed after the two lab sessions, but we will change this as described above, so that the participants do not have to come in again to collect their money or course credit.

## Analysis

We aim to conduct all analysis as described in Westbrook et al. [-@Westbrook2013], but the level of detail was not always sufficient, so there might be deviations regarding data cleaning and degrees of freedom.
The performance measure *d'* will be computed as the difference of the *z*-transformed hit rate and the *z*-transformed false alarm rate [@Macmillan1990].
Reaction time (RT) data will be trimmed by excluding all trials with responses faster than 100 ms, as the relevant cognitive processes cannot have been completed before [@Whelan2008; @Berger2021].
Aggregated RT values will be described using the median and the median of absolute deviation (MAD) as robust estimates of center and variability, respectively [@Lachaud2011].
Error- and post-error trials will be excluded in repeated measures analyses of variance (rmANOVA) and controlled for in multi-level-model (MLM), because RT on the latter is longer due to more cautious behavior [@Dutilh2012; @Houtman2012].
To test our hypotheses, we will perform a series of rmANOVAs and an MLM with orthogonal sum-to-zero contrasts in order to meaningfully interpret results [@Singmann2019]. 
Declining performance will be investigated by calculating an rmANOVA with three paired contrasts comparing *d'* between two levels of 2-, 3-, and 4-back at a time.
Another rmANOVA with three paired contrasts will be computed to compare the mean RT between two levels of 2-, 3-, and 4-back at a time.
To investigate changes in NASA-TLX ratings, six rmANOVAs will be computed, one for each NASA-TLX subscale, and each with six paired contrasts comparing the ratings between two levels of 1-, 2-, 3-, and 4-back at a time.
For each ED round, SVs will be calculated by adding or subtracting 0.015625 from the last monetary value of the flexible level, depending on the participant's last choice.
Then, these final monetary values will be divided by 2€, and the SV of each level per participant will be computed by averaging all final values of each level, regardless of whether it was fixed or flexible.
An rmANOVA with six paired contrasts will be computed, comparing the SVs between two levels of 1-, 2-, 3-, and 4-back at a time.
Tukey method will be used for the paired contrasts of each rmANOVA, including *p*-value adjustment.

To determine the influence of task performance in the association of SVs and n-back level, we will set up a MLM using the $lmerTest$ package. 
We apply restricted maximum likelihood (REML) to fit the model.
First, we calculate the intraclass correlation (ICC) on the basis of the null model.
Second, we estimate a random slopes model of SVs including n-back load level as level-1-predictor and, additionally, NFC as level-2-predictor. 
Within the model, we control for *d'*, RT, correct and post-correct trials.
$$
SV \sim level\ * NFC + d' + RT + correct + postcorrect + (level|subject)
$$
Level-1-predictors are centered within cluster, whereas the level-2-predictor is centered at the grand mean as recommended by Enders & Tofighi [-@Enders2007]. 
We visually inspect the residuals of the final model.
The approximately normal distribution indicates no evidence to perform model criticism.  
Third, we perform a simple slopes analysis with n-back level as predictor and NFC as moderator. 
To evaluate the moderating effect, we calculate the Johnson-Neyman interval.

To ensure the validity of the multi-level model, we will conduct a specification curve analysis [@Simonsohn2020], which will include 63 possible preprocessing pipelines of the RT data.
These pipelines specify which transformation was applied (none, log, inverse, or square-root), which outliers were excluded (none, 2, 2.5, or 3 $MAD$ from the median, RTs below 100 or 200 ms), and across which dimensions the transformations and exclusions were applied (across/within subjects and across/within n-back levels).
The multi-level model will be run with each of the 63 pipelines, which will also include our main pipeline (untransformed data, exclusion of RTs below 100 ms).
The ratio of pipelines that lead to significant versus non-significant effects will provide an indication of how robust the effect actually is.

@@

The results of each analysis will be assessed on the basis of both *p*-value and the Bayes factor $BF10$.

## Power analysis

A sample size analysis with GPower [@Faul2007; @Faul2009], using the pilot RT data, indicated that we should collect data from at least 53 participants, assuming $f=0.20$, $\alpha=.05$, and $\beta=.95$.

@@

>>>>

---
title             : "When easy is not preferred: An effort discounting paradigm for estimating subjective values of tasks"
shorttitle        : "The CERED paradigm for estimating subjective values"
author: 
  - name          : "Josephine Zerna"
    orcid         : "0000-0003-2892-884X"
    affiliation   : "1,*"
    corresponding : yes
    address       : "Zellescher Weg 17, 01069 Dresden, Germany"
    email         : "josephine.zerna@tu-dresden.de"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Methodology
      - Formal analysis
      - Investigation
      - Project administration
      - Software
      - Visualization
      - Writing - original draft preparation
      - Writing - review & editing
  - name          : "Christoph Scheffel"
    orcid         : "0000-0001-5963-9229"
    affiliation   : "1,*"
    role:
      - Conceptualization
      - Methodology
      - Funding acquisition
      - Investigation
      - Project administration
      - Software
      - Writing - review & editing
  - name          : "Corinna Kührt"
    affiliation   : "1"
    role:
      - Formal analysis
      - Writing - original draft preparation
      - Writing - review & editing
      - Visualization
affiliation:
  - id            : "1"
    institution   : "Faculty of Psychology, Technische Universität Dresden, 01069 Dresden, Germany"
authornote: |
  * Shared first authorship
abstract: |
  When individuals set goals, they consider the subjective value (SV) of both the anticipated reward and the required effort, a trade-off that is of great interest to psychological research.
  However, the SV of effort is highly individual, and previous quantification approaches have had two crucial limitations: They presumed a unidirectional relationship between objective task load and the SV of effort, and as a consequence, they could only quantify the SVs of tasks with an objective order of task load.
  One of these approaches is the Cognitive Effort Discounting paradigm by Westbrook et al. (2013).
  We aim to replicate their analysis with our adaptation, the Cognitive and Emotion Regulation Effort Discounting (CERED) paradigm.
  We argue that the CERED paradigm allows two crucial things: Quantifying SVs without assuming that the easiest level is preferred, and quantifying SVs for tasks with no objective order of task load.

@@

The paradigm was written and presented using *Psychopy* [@Peirce2019]. We used *R Studio* [@RCT2020; @RStudioTeam2020] with the main packages *afex* [@Singmann2021] and *bayestestR* [@Makowski2019] for all our analyses.

## Ethics information

## Pilot data

The sample of the pilot study consisted of $N=$ `r length(datalist_ED)` participants (`r round((table(data_quest$gender)[2]/length(datalist_ED))*100, digits = 1)`% female, $M=$`r round(mean(data_quest$age, na.rm = TRUE), digits = 1)`($SD=$`r round(sd(data_quest$age, na.rm = TRUE), digits = 1)`) years old).

## Design

Healthy participants aged 18 to 30 years will be recruited using the software *ORSEE* [@Greiner2015].
Participants will fill out the personality questionnaires online and then visit the lab for two sessions one week apart.
NFC will be assessed using the 16-item short form of the Need for Cognition Scale [@Bless1994; @Cacioppo1984].
Responses to each item (e.g., "Thinking is not my idea of fun", recoded) will be recorded on a 7-point Likert scale.
The NFC scale shows comparably high internal consistency [Cronbach's $\alpha$ \> .80; @Bless1994; @Fleischhauer2010].
Several other personality questionnaires will be used in this study but are the topic of the Registered Report for the second lab session.
A full list of measures can be found in our [Github repository](https://github.com/ChScheffel/CERED).
In the first session, participants provide informed consent and demographic data before completing the computer-based paradigm.
The paradigm starts with the n-back levels one to four, presented sequentially with two runs per level, consisting of 64 consonants (16 targets, 48 non-targets) per run.
The levels are referred to by color (1-back black, 2-back red, 3-back blue, 4-back green) to avoid anchor effects in the ED procedure.
To assess perceived task load, we will use the 6-item NASA Task Load Index (NASA-TLX) [@Hart1988], where participants evaluate their subjective perception of mental load, physical load, effort, frustration, performance, and time pressure during the task on a 20-point scale.
After each level, participants fill out the NASA-TLX on a tablet.
Then, they complete the ED procedure on screen, where each possible pairing of the four n-back levels is presented in a randomized order.
Participants are instructed to decide as realistically as possible, because one of their choices from the last iteration steps will be randomly chosen for one final run of n-back.
This is only done to incentivise truthful behavior in the ED procedure, so the n-back data of this part will not be analyzed.
The second session consists of an emotion regulation task with negative pictures and the instruction to suppress facial reactions, detach cognitively from the picture content, and distract oneself, respectively.
The paradigm follows the same structure of task and ED procedure, but participants can decide which strategy they want to reapply in the last block.
Participants will receive 30€ in total or course credit for participation.
Study data will be collected and managed using REDCap electronic data capture tools hosted at Technische Universität Dresden [@Harris2009; @Harris2019].

## Sampling plan

A sample size analysis with G*Power [@Faul2007; @Faul2009], based on the results of the ANOVA of Westbrook et al. [-@Westbrook2013] which showed an increase in reaction time with higher n-back levels, indicated that we should collect data from at least 53 participants, assuming $\eta2=0.04$, $\alpha=.05$, and $\beta=.95$.
The power analyses of all other hypotheses yielded smaller necessary sample sizes.
To account for technical errors and exclusions of physiological data of the second lab session due to excessive noise, we aim to collect data of 60 to 70 participants.

## Analysis plan

Data collection and analysis will not be performed blind to the conditions of the experiments.
We aim to conduct all analysis as described in Westbrook et al. [-@Westbrook2013], but the level of detail was not always sufficient, so there might be deviations regarding data cleaning and degrees of freedom.
The performance measure *d'* will be computed as the difference of the *z*-transformed hit rate and the *z*-transformed false alarm rate [@Macmillan1990].
Reaction time (RT) data will be trimmed by excluding all trials with responses faster than 100 ms, as the relevant cognitive processes cannot have been completed before [@Whelan2008; @Berger2021].
Aggregated RT values will be described using the median and the median of absolute deviation (MAD) as robust estimates of center and variability, respectively [@Lachaud2011].
Error- and post-error trials will be excluded in repeated measures analyses of variance (rmANOVA) and controlled for in multi-level-model (MLM), because RT on the latter is longer due to more cautious behavior [@Dutilh2012; @Houtman2012].
To test our hypotheses, we will perform a series of rmANOVAs and an MLM with orthogonal sum-to-zero contrasts in order to meaningfully interpret results [@Singmann2019]. 
Declining performance will be investigated by calculating an rmANOVA with three paired contrasts comparing *d'* between two levels of 2-, 3-, and 4-back at a time.
Another rmANOVA with three paired contrasts will be computed to compare the mean RT between two levels of 2-, 3-, and 4-back at a time.
To investigate changes in NASA-TLX ratings, six rmANOVAs will be computed, one for each NASA-TLX subscale, and each with six paired contrasts comparing the ratings between two levels of 1-, 2-, 3-, and 4-back at a time.
For each ED round, SVs will be calculated by adding or subtracting 0.015625 from the last monetary value of the flexible level, depending on the participant's last choice.
Then, these final monetary values will be divided by 2€, and the SV of each level per participant will be computed by averaging all final values of each level, regardless of whether it was fixed or flexible.
An rmANOVA with six paired contrasts will be computed, comparing the SVs between two levels of 1-, 2-, 3-, and 4-back at a time.
Tukey method will be used for the paired contrasts of each rmANOVA, including *p*-value adjustment.

To determine the influence of task performance in the association of SVs and n-back level, we will set up a MLM using the $lmerTest$ package. 
We will apply restricted maximum likelihood (REML) to fit the model.
First, we will calculate the intraclass correlation (ICC) on the basis of the null model.
Second, we will estimate a random slopes model of SVs including n-back load level as level-1-predictor and, additionally, NFC as level-2-predictor. 
Within the model, we will control for *d'*, RT, correct, and post-correct trials.
$$
SV \sim level\ * NFC + d' + RT + correct + postcorrect + (level|subject)
$$
Level-1-predictors will be centered within cluster, whereas the level-2-predictor will be centered at the grand mean as recommended by Enders & Tofighi [-@Enders2007]. 
We will visually inspect the residuals of the final model.
The approximately normal distribution indicates no evidence to perform model criticism.  
Third, we will perform a simple slopes analysis with n-back level as predictor and NFC as moderator. 
To evaluate the moderating effect, we will calculate the Johnson-Neyman interval.

To ensure the validity of the MLM, we will conduct a specification curve analysis [@Simonsohn2020], which will include 63 possible preprocessing pipelines of the RT data.
These pipelines specify which transformation was applied (none, log, inverse, or square-root), which outliers were excluded (none, 2, 2.5, or 3 $MAD$ from the median, RTs below 100 or 200 ms), and across which dimensions the transformations and exclusions were applied (across/within subjects and across/within n-back levels).
The MLM will be run with each of the 63 pipelines, which will also include our main pipeline (untransformed data, exclusion of RTs below 100 ms).
The ratio of pipelines that lead to significant versus non-significant effects will provide an indication of how robust the effect actually is.

@@

The results of each analysis will be assessed on the basis of both *p*-value and the Bayes factor $BF10$, calculated using the *BayesFactor* package [@Morey2021].

## Data availability

The data of this study can be downloaded from osf.io/vnj8x/.

## Code availability

The paradigm code, as well as the R Markdown file used to analyze the data and write this document is available at our [Github repository](https://github.com/ChScheffel/CERED).

@@

This research was partly funded by the German Research Foundation (DFG) as part of the Collaborative Research Center (CRC) 940.
The funders have/had no role in study design, data collection and analysis, decision to publish or preparation of the manuscript.

# Author Contributions

# Competing Interests

The authors declare no competing interests.

# Figure Captions

# Design Table

# Supplement
